---
title: "ANA 515 Assignment 3"
author: "Mounika Gangishetty"
output:
  pdf_document: default
  html_document:
    theme:
      bootswatch: superhero
 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE}
library(dplyr)
library(readxl)
library(ggplot2)
library(openxlsx)
```

# Reading the GRAIN data from downloaded excel

```{r, echo = TRUE}
#  reading excel from local drive

setwd("/Users/vamshinaidi/Documents/ANA-515/PracticumProject")
excel_file  <- 'GRAINLandJan2012.xlsx'

# reading data from sheet1

sheet1 <- read_excel(excel_file,sheet="Sheet1")
head(sheet1)

# reading data from sheet2

sheet2 <- read_excel(excel_file,sheet="Sheet2")
head(sheet2)

```

# Using unique() function to find the distinct values in the "Landgrabbed" column, i.e distinct country names from sheet1.


This data is about sales of vast amounts of agricultural land in less developed countries.
The dataset has  information about `r length(unique(sheet1$Landgrabbed))` countries agricultural land sales data.
The datset has  few null values for columns like 'Projected investment' ,'Year'  and invalid data for columns
like Status of Deal' and 'Year'.
We are going to remove these outliers and anamolies from the dataset.



# Removing missing values, null, NA and invalid data from the datset

```{r, echo = TRUE}

# Removing the sales data in the sheets where projected investment amount is null,
# Hectares is null and  invalid years

filter_sheet1 <- sheet1 %>% 
  filter(!is.na(`Projected investment`) & !is.null(Year) & !is.na(Year) & !is.na(`Hectares`))

head(filter_sheet1)

filter_sheet2 <- sheet2 %>% 
  filter(!is.na(`Projected investment`) & !is.null(Year) & !is.na(Year) & !is.na(`Hectares`))


head(filter_sheet2)


```


# Correcting the misspelled data for data consistency 

```{r, echo = TRUE}
#
# Spell correction in column "Status of deal"
#
# Printing the values before changes
#
head(filter_sheet1$'Status of deal')

head(filter_sheet2$'Status of deal')

# 
# we are using mutate function to handle misspelled data ,replacing "Don" with correct spell "Done"
#  in the Status column and remaining we are not changing by passing the parameter TRUE in mutate function


filter_sheet1 <- filter_sheet1 %>%
  mutate(`Status of deal` = case_when(
    `Status of deal` == "Don" ~ "Done",
    TRUE ~ `Status of deal`  
  ))


filter_sheet2 <- filter_sheet2 %>%
  mutate(`Status of deal` = case_when(
    `Status of deal` == "Don" ~ "Done",
    TRUE ~ `Status of deal` 
  ))

# Data after spell correction

head(filter_sheet1$'Status of deal')


head(filter_sheet2$'Status of deal')


```

# Combining the two data sheets into one

```{r, echo = TRUE}

# Combining the data from sheet1 and sheet2 into single file.

combined_data <- rbind(filter_sheet1,filter_sheet2)

head(combined_data)

```

# Sactter plot and histogram to understand  hectares of land sales in less devloped countries among many years.

```{r, echo = TRUE}

#
# scatterplot determining  "Hectares of Land" vs. "Projected investment"
# converting the values to numeric before proceeding to plot to avoid  "need finite 'xlim' values" error
#

combined_data$Hectares <- as.numeric(combined_data$Hectares)
combined_data$`Projected investment` <- as.numeric(combined_data$`Projected investment`)

plot(combined_data$Hectares, 
     combined_data$`Projected investment`, 
     main = "Scatterplot of Hectares vs. Projected Investment", 
     xlab = "Hectares", 
     ylab = "Projected Investment", 
     pch = 19, 
     col = "blue")

#
# A 2d histogram represents the hectares of land sales year wise
#

ggplot(combined_data, aes(x = Year, y = Hectares)) +
  geom_bin2d(bins = 30) +
  labs(
    title = "2D Histogram",
    x = "Projected Year",
    y = "Land area in Hectares"
  )


```

# Writing the cleaned data set named "filter_sheet1", "filter_sheet2" into an new  excel file

```{r, echo = TRUE}
wb <- createWorkbook()

# We have filtered data sheets  "filter_sheet1", "filter_sheet2"

# Adding each filtered sheet into the new Excel file
addWorksheet(wb, sheetName = "Sheet1")
writeData(wb, sheet = "Sheet1", filter_sheet1)

addWorksheet(wb, sheetName = "Sheet2")
writeData(wb, sheet = "Sheet2", filter_sheet2)

# Save the Excel workbook to a file
saveWorkbook(wb, file = "filtered_data_Mounika.xlsx")

```